---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r set-options, echo=FALSE, cache=FALSE}
library(sqlscore)
library(knitr)

options(width=80)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  tidy = TRUE,
  size="small",
  linewidth=80
)

# From https://github.com/yihui/knitr-examples/blob/master/077-wrap-output.Rmd
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})
```

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/sqlscore)](http://cran.r-project.org/web/packages/sqlscore)
[![Downloads](http://cranlogs.r-pkg.org/badges/sqlscore)](http://cran.rstudio.com/package=sqlscore)
[![License](http://img.shields.io/:license-mit-blue.svg?style=flat)](http://badges.mit-license.org)

# sqlscore
R utilities to score GLMs and related models in SQL.

---

The sqlscore package provides utilities for generating sql queries
(particularly CREATE TABLE statements) from R model objects. The most important
use case is generating SQL to score a GLM or related model represented as an R
object, which is particularly important when the amount of scoring data is very large.
Scoring in the database avoids having to read all of the data into memory (and
may also avoid other types of I/O overhead).

The SQL-generating functions in sqlscore handle various formula operators, and
also take care of wrapping the model's linear predictor in the appropriate link
function. If needed, you can also specify a custom link function.

## Usage:
The SQL-generating functions create\_statement and select\_statement do what their
names suggest and generate CREATE TABLE and SELECT statements for model scoring.

If, for example, you have a database table of iris measurements, you can model the
sepal length and generate predictions as follows:
```
mod <- glm(Sepal.Length ~ Sepal.Width + Petal.Length + Petal.Width + Species,
           data=datasets::iris)

create_statement(mod, src_table="iris", dest_table="iris_scores")  
```

```{r, echo = FALSE}
mod <- glm(Sepal.Length ~ Sepal.Width + Petal.Length + Petal.Width + Species,
           data=datasets::iris)

create_statement(mod, src_table="iris", dest_table="iris_scores")  
```

To get a select statement that's not wrapped in a CREATE TABLE (so that, e.g.,
you can add your own database-specific pieces of SQL), use select_statement:
```
mod <- glm(Sepal.Length ~ Sepal.Width + Petal.Length + Petal.Width + Species,
           data=datasets::iris)

create_statement(mod, src_table="iris", dest_table="iris_scores")  
```

```{r, echo = FALSE}
mod <- glm(Sepal.Length ~ Sepal.Width + Petal.Length + Petal.Width + Species,
           data=datasets::iris)

create_statement(mod, src_table="iris", dest_table="iris_scores")  
```

Helper functions include linpred(), which generates an R call object representing
the linear predictor, and score_expression, an S3 generic that handles wrapping
the linear predictor in the response function.

## Supported model types:
Specific packages and models that are known to work include: glm and lm from
package:stats, cv.glmnet from package:glmnet, glmboost from package:mboost,
and bayesglm from package:arm.

Default S3 methods are for objects structured like those of class "glm", so
models not listed here may work if they resemble those objects, but are not 
guaranteed to.

## Installation:
Install the released version from CRAN:
```
install.packages('sqlscore')
```

Install the dev version from github:
```
install.packages("devtools")
devtools::install_github("wwbrannon/sqlscore")
```
